A Data-Driven Framework for Neural Field Modelling

### 2010-10-28 ltx/first_order_framework/LaTeX_Manuscript.tex
got rid of fmax

The firing rate of the presynaptic neurons is related to the post-synaptic membrane potential by the sigmoidal activation function 
\begin{equation}
	\label{ActivationFunction} f\left( v\left( \mathbf{r}', t \right) \right) = \frac{f_{max}}{1 + \exp \left( \varsigma \left( v_0 - v\left(\mathbf{r}',t\right) \right) \right)}. 
\end{equation}
The parameters $f_{max}$ and $v_0$ describe the maximum firing rate and firing threshold of the neural populations respectively. The parameter $\varsigma$ governs the slope of the sigmoid. By substituting equation~\ref{RateBasedInteractions} into equation~\ref{SpikesToPotential} we get the spatiotemporal model 
\begin{equation}
	\label{FullDoubleIntModel} v\left(\mathbf{r},t\right) =
	\int_{-\infty}^t 
	h\left(t - t'\right) \int_\Omega
	w\left(\mathbf{r},\mathbf{r}'\right) 
	f\left( v\left( \mathbf{r}',t' \right)\right)
	\textrm{d}\mathbf{r}'\textrm{d}t'.
\end{equation}
To arrive at the final form of the model, we shall express the synaptic response kernel as a Green's function 
\begin{equation}
	\label{GreensFuncDef} \textrm{D}h\left( t \right) = \delta \left( t \right), 
\end{equation}
where $\textrm{d}=\frac{d}{\textrm{d}t} + \zeta$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function giving 
\begin{equation}
	\label{FinalFormContinuous} 
	\frac{\textrm{d}v\left( \mathbf{r},t \right)}{\textrm{d}t} + \zeta v\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)\textrm{d}\mathbf{r}'}. 
\end{equation}
To arrive at the integro-difference equation (IDE) form of the model, we discretise time using a first-order Euler method (see the S1 for derivation) giving 
\begin{equation}
	\label{DiscreteTimeModel} 
	v_{t+T_s}\left(\mathbf{r}\right) = 
	\xi v_t\left(\mathbf{r}\right) + 
	T_s \int_\Omega { 
	    w\left(\mathbf{r},\mathbf{r}'\right)
	    f\left(v_t\left(\mathbf{r}'\right)\right) 
	\textrm{d}\mathbf{r}'} 
	+ e_t\left(\mathbf{r}\right), 
\end{equation}
where $T_s$ is the time step, $\xi = 1-T_s\zeta $ and $e_t(\mathbf{r})$ is an $i.i.d.$ disturbance such that $e_t(\mathbf{r})\sim\mathcal{GP}(\mathbf 0,\gamma(\mathbf{r}-\mathbf{r}'))$. Here $\mathcal{GP}(\mathbf 0,\gamma(\mathbf{r}-\mathbf{r}'))$ denotes a zero mean spatial Gaussian process with covariance function $\gamma(\mathbf{r}-\mathbf{r}')$~\cite{Rasmussen2005}. The disturbance is added to account for model uncertainty and unmodelled inputs. To simplify the notation, the index of the future time sample, $t+T_s$, shall be referred to as $t+1$ throughout the rest of the paper.

@@

\begin{table}[!ht]
\caption{
\bf{Parameters}}
\begin{tabular}{|c|l|c|l|}
	\hline
	\textbf{Symbol} & \textbf{Quantity} & \textbf{Value} & \textbf{Units}\\
	\hline
	\multicolumn{4}{|c|}{\emph{Domain and indices}}\\
	\hline
	$\Delta$ & spatial discretisation step & 0.5 & mm \\
	$T_s$ & time step & $0.001$ & s\\
	$T$ & number of time steps used in estimation & 400 & n.a.\\
	\hline 
\multicolumn{4}{|c|}{\emph{Model}}\\
	\hline
	$\tau$ & synaptic time constant & 0.01 & s$^{-1}$ \\
	$f_{max}$ & maximal firing rate & 10 & spikes.s$^{-1}$\\
	$\varsigma$ & slope of sigmoidal activation function & 0.8 & mV$^{-1}$\\
	$v_0$ & firing threshold & 2 & mV \\
	$\boldsymbol{\theta}$ & vector of connectivity kernel parameters & $\left[\begin{array}{ccc}
	10 &-8 &0.5
	\end{array}
	\right]^{\top}$ & mV.spike$^{-1}$\\
	$\sigma_{\psi_{1}}, \sigma_{\psi_{2}}, \sigma_{\psi_{3}}$ & connectivity kernel width parameters & 1.8, 2.4, 6 & mm\\
	$N$ & number of sensors & 196 & n.a.\\ 
	$\Delta_y$ & distance between adjacent sensors & 1.5 & mm\\
	$\sigma_{m}$ & sensor width parameter & 0.9 & mm \\
	$\Sigma_{\varepsilon}$ & observation noise variance & $0.1 \times I_{N}$ & mm$^2$ \\
	$\sigma_{\gamma}$& disturbance spatial covariance width & 1.3 & mm\\
	$\sigma_{d}$ & disturbance variance & 0.1 & mV \\
	\hline 
	\multicolumn{4}{|c|}{\emph{Reduced Model}}\\
	\hline
	$L$ & number of basis functions& 81 & n.a.\\
	$\Delta_{\phi}$ & distance between field basis functions & 2.5 & mm \\
	$\sigma^2_{\phi}$ & Spatial variance of Gaussian in spatial domain & 2.5 & mm$^2$\\
	\hline 
	\multicolumn{4}{|c|}{\emph{Frequency Analysis}}\\
	\hline
	$\rho_y$ & sensor over-sampling parameter & 1 & n.a.\\
	$\rho_\phi$ & field basis function over-sampling parameter & 1 & n.a.\\
	\hline 
	\multicolumn{4}{|c|}{\emph{Estimation}}\\
	\hline
	$\alpha$& determines the spread of sigma points & 0.001 & n.a.\\
	$\beta$& prior knowledge of the sigma points distribution & 2 & n.a.\\
	$\kappa$& scaling parameter $(3-L)$& -78 & n.a.\\
	$\lambda$& scaling parameter & -80.99 & n.a.\\
	\hline
\end{tabular}
\begin{flushleft}Parameters for the neural field model, the model reduction, the frequency analysis and the estimation procedure.
\end{flushleft}
\label{tab:Parameters}
\end{table}

>>>>

The firing rate of the presynaptic neurons is related to the post-synaptic membrane potential by the sigmoidal activation function 
\begin{equation}
	\label{ActivationFunction} f\left( v\left( \mathbf{r}', t \right) \right) = \frac{1}{1 + \exp \left( \varsigma \left( v_0 - v\left(\mathbf{r}',t\right) \right) \right)}. 
\end{equation}
The parameter $v_0$ describes the  firing threshold of the neural populations. The parameter $\varsigma$ governs the slope of the sigmoid. By substituting equation~\ref{RateBasedInteractions} into equation~\ref{SpikesToPotential} we get the spatiotemporal model 
\begin{equation}
	\label{FullDoubleIntModel} v\left(\mathbf{r},t\right) =
	\int_{-\infty}^t 
	h\left(t - t'\right) \int_\Omega
	w\left(\mathbf{r},\mathbf{r}'\right) 
	f\left( v\left( \mathbf{r}',t' \right)\right)
	\textrm{d}\mathbf{r}'\textrm{d}t'.
\end{equation}
To arrive at the final form of the model, we shall express the synaptic response kernel as a Green's function 
\begin{equation}
	\label{GreensFuncDef} \textrm{D}h\left( t \right) = \delta \left( t \right), 
\end{equation}
where $\textrm{d}=\frac{d}{\textrm{d}t} + \zeta$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function giving 
\begin{equation}
	\label{FinalFormContinuous} 
	\frac{\textrm{d}v\left( \mathbf{r},t \right)}{\textrm{d}t} + \zeta v\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)\textrm{d}\mathbf{r}'}. 
\end{equation}
To arrive at the integro-difference equation (IDE) form of the model, we discretise time using a first-order Euler method (see the S1 for derivation) giving 
\begin{equation}
	\label{DiscreteTimeModel} 
	v_{t+T_s}\left(\mathbf{r}\right) = 
	\xi v_t\left(\mathbf{r}\right) + 
	T_s \int_\Omega { 
	    w\left(\mathbf{r},\mathbf{r}'\right)
	    f\left(v_t\left(\mathbf{r}'\right)\right) 
	\textrm{d}\mathbf{r}'} 
	+ e_t\left(\mathbf{r}\right), 
\end{equation}
where $T_s$ is the time step, $\xi = 1-T_s\zeta $ and $e_t(\mathbf{r})$ is an $i.i.d.$ disturbance such that $e_t(\mathbf{r})\sim\mathcal{GP}(\mathbf 0,\gamma(\mathbf{r}-\mathbf{r}'))$. Here $\mathcal{GP}(\mathbf 0,\gamma(\mathbf{r}-\mathbf{r}'))$ denotes a zero mean spatial Gaussian process with covariance function $\gamma(\mathbf{r}-\mathbf{r}')$~\cite{Rasmussen2005}. The disturbance is added to account for model uncertainty and unmodelled inputs. To simplify the notation, the index of the future time sample, $t+T_s$, shall be referred to as $t+1$ throughout the rest of the paper.

@@

\begin{table}[!ht]
\caption{
\bf{Parameters}}
\begin{tabular}{|c|l|c|l|}
	\hline
	\textbf{Symbol} & \textbf{Quantity} & \textbf{Value} & \textbf{Units}\\
	\hline
	\multicolumn{4}{|c|}{\emph{Domain and indices}}\\
	\hline
	$\Delta$ & spatial discretisation step & 0.5 & mm \\
	$T_s$ & time step & $0.001$ & s\\
	$T$ & number of time steps used in estimation & 400 & n.a.\\
	\hline 
\multicolumn{4}{|c|}{\emph{Model}}\\
	\hline
	$\tau$ & synaptic time constant & 0.01 & s$^{-1}$ \\
% 	$f_{max}$ & maximal firing rate & 10 & spikes.s$^{-1}$\\
	$\varsigma$ & slope of sigmoidal activation function & 0.8 & mV$^{-1}$\\
	$v_0$ & firing threshold & 2 & mV \\
	$\boldsymbol{\theta}$ & vector of connectivity kernel parameters & $\left[\begin{array}{ccc}
	10 &-8 &0.5
	\end{array}
	\right]^{\top}$ & mV.spike$^{-1}$\\
	$\sigma_{\psi_{1}}, \sigma_{\psi_{2}}, \sigma_{\psi_{3}}$ & connectivity kernel width parameters & 1.8, 2.4, 6 & mm\\
	$N$ & number of sensors & 196 & n.a.\\ 
	$\Delta_y$ & distance between adjacent sensors & 1.5 & mm\\
	$\sigma_{m}$ & sensor width parameter & 0.9 & mm \\
	$\Sigma_{\varepsilon}$ & observation noise variance & $0.1 \times I_{N}$ & mm$^2$ \\
	$\sigma_{\gamma}$& disturbance spatial covariance width & 1.3 & mm\\
	$\sigma_{d}$ & disturbance variance & 0.1 & mV \\
	\hline 
	\multicolumn{4}{|c|}{\emph{Reduced Model}}\\
	\hline
	$L$ & number of basis functions& 81 & n.a.\\
	$\Delta_{\phi}$ & distance between field basis functions & 2.5 & mm \\
	$\sigma^2_{\phi}$ & Spatial variance of Gaussian in spatial domain & 2.5 & mm$^2$\\
	\hline 
	\multicolumn{4}{|c|}{\emph{Frequency Analysis}}\\
	\hline
	$\rho_y$ & sensor over-sampling parameter & 1 & n.a.\\
	$\rho_\phi$ & field basis function over-sampling parameter & 1 & n.a.\\
	\hline 
	\multicolumn{4}{|c|}{\emph{Estimation}}\\
	\hline
	$\alpha$& determines the spread of sigma points & 0.001 & n.a.\\
	$\beta$& prior knowledge of the sigma points distribution & 2 & n.a.\\
	$\kappa$& scaling parameter $(3-L)$& -78 & n.a.\\
	$\lambda$& scaling parameter & -80.99 & n.a.\\
	\hline
\end{tabular}
\begin{flushleft}Parameters for the neural field model, the model reduction, the frequency analysis and the estimation procedure.
\end{flushleft}
\label{tab:Parameters}
\end{table}
